---
title: "Project 2"
author: "Arthur Gogohia, Dion Refiano, Konstantin Shuxtelinsky"
date: "26/05/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd('/Users/Kostja/Desktop/Master/Sem 2 (18 SoSe)/Data Visualization/Tasks/DV_Project_2/data')
```

## Data introduction
As required, this task was an open one, so the students had to choose a specific topic on their own. Our Group did choose a dataset we found on https://labrosa.ee.columbia.edu/millionsong/pages/getting-dataset#subset. This subset Contains 10k Music files and is around 2GB big. The actual dataset is about 300GB big and has arround 1 MIllion entries, in this case songs. Besids the Analysis the dataset contains also some Metadata, like Author, produced year etc. The actual Provider of this data set is THE ECHO NEST (http://the.echonest.com). As provided by the information about the dataset, it is a result of an collaboration between THE ECHO NEST and LabROSA (https://labrosa.ee.columbia.edu). 

## Handle the downloaded data
After downloading and unzipping the data, one can see two different dolders, one 'data' containing several other folders 'A' and 'B' which contain songfiles in HDF5 (Hirarchical Data Format 5) format. This is a format that is still used in scientific applications and using and more saving big datasets.

```{r import}
#install.packages("wordcloud") # word-cloud generator
#install.packages("RColorBrewer") # color palettes

require("wordcloud")
library("RColorBrewer")

```


```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
